---
title: "EDA5-shuting"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidyr)
library(dplyr)
library(ggplot2)
library(plotly)
# open XQuartz
library(TSdist)
library(lubridate)
library(tibble)
library(factoextra)
library(rbenchmark)
library(autoencoder)
```

## import data
```{r}
actual_daily <- read_csv("~/Desktop/BU-2021fall/partner project/fidelity-github/11.14备份/lst/data/actual_daily.csv")
actual_daily$date <- as.character(actual_daily$date)

execution_daily <- read.csv("~/Desktop/BU-2021fall/partner project/fidelity-github/11.14备份/lst/data/execution_daily.csv")

efficiency_daily <- inner_join(actual_daily,execution_daily,by=c("NEWID","date"))
efficiency_daily$efficiency <- log(efficiency_daily$sum.y/efficiency_daily$sum.x)
efficiency_daily$efficiency_exp <- efficiency_daily$sum.y/efficiency_daily$sum.x
write.csv(efficiency_daily,file="~/Desktop/efficiency_daily.csv")
```

## compute effi & ac distance of one warehouse
```{r}
##test, choose ID=9, date from 650 to 732
a <- actual_daily[actual_daily$NEWID==9,5]
a1 <- a[650:732,]
eff <- efficiency_daily[efficiency_daily$NEWID==9,9]
eff <- as.matrix(eff)
eff[is.nan(eff)]<-0
eff1 <- eff[650:732,]
# da=dist(a1,eff1, method="TSDistances", distance="dtw", sigma=10)
# dim(da)

DTWDistance(a1,eff1, window.type="sakoechiba", window.size=7)

  actual <- actual_daily[c(650:732),5]
  effient <- efficiency_daily[c(650:732),9]
  eff <- as.matrix(eff)
  eff[is.nan(eff)]<-0
  DTWDistance(a1,eff1, window.type="sakoechiba", window.size=7)
```


## loop for 320 warehouses to compute efficiency & actual credit distance
```{r,warning=FALSE}
actual_daily$date <-as.Date(actual_daily$date)
efficiency_daily$date <-as.Date(efficiency_daily$date)
dis <- rep(NA,320)

effi_actual_distance <- function(begin,end){
#  begin <- "2021-03-29"
#  end <- "2021-07-04"
  actual_0 <- actual_daily[actual_daily$date >=begin & actual_daily$date<= end,]
  effient_0 <- efficiency_daily[efficiency_daily$date >=begin & efficiency_daily$date<= end,] 
  for (i in 1:320){
#    i <- 190
  actual <- actual_0[actual_0$NEWID==i,5] 
  effient <- effient_0[effient_0$NEWID==i,9]
  effient <- as.matrix(effient)
  effient[is.nan(effient)]<-0
  effient[!is.finite(effient)]<-0
  dis[i] <- DTWDistance(actual,effient, window.type="sakoechiba", window.size=7)
  }
return(dis)
}

dis_final <- effi_actual_distance("2021-03-29","2021-07-04")
dis_final <- data.frame(NEWID=c(1:320),dis_final)
#length(which(dis_final != 0))
```

## group warehouses based on distance between effiency and actual credit
```{r}
##plot
scatter <- ggplot(dis_final)+
  geom_point(aes(x=NEWID,y=dis_final))+
  ylim(0,15000)
ggplotly(scatter)
```
From scatter plot, NEWID=9 has extermely high distance, over 14000, which means it's efficiency pattern is extremely different with its actual credit pattern from 2021-03-29 to 2021-07-04.


```{r}
## drop NEWID = 9
df <- dis_final[-9,]
df.scaled <- scale(df[,-1])
## choose best K
fviz_nbclust(df,kmeans, method = "wss") +
geom_vline(xintercept = 4, linetype = 2)   ## Best K = 4
## Compute k-means with k = 4
set.seed(1)
kmdis <- kmeans(df.scaled, centers = 4, nstart = 25)
print(kmdis)
## Visualize k-means
fviz_cluster(kmdis, df, geom="text") ##use it
fviz_cluster(kmdis, df, ellipse.type = "norm")

## drop NEWID = 274
df1 <- df[-273,]
df1.scaled <- scale(df1[,-1])
set.seed(111)
kmdis1 <- kmeans(df1.scaled, centers = 3, nstart = 25)
fviz_cluster(kmdis1, df1, geom="text")

## add cluster info to original data
table <- cbind(df, cluster = kmdis$cluster)
```

TIPS:
As the final result of k-means clustering result is sensitive to the random starting assignments, we specify nstart = 25. This means that R will try 25 different random starting assignments and then select the best results corresponding to the one with the lowest within cluster variation.





# try to compare exe & actual credit

## loop for 320 warehouses to compute execution & actual credit distance
```{r,warning=FALSE}
actual_daily$date <-as.Date(actual_daily$date)
execution_daily$date <-as.Date(execution_daily$date)

## Define function to compute distance in a time period
exe_actual_distance <- function(begin,end,take_log=FALSE){
  dis_ac_exe <- rep(NA,320)
  acct_ave <- rep(NA,320)
  exec_ave <- rep(NA,320)
  actual_0 <- actual_daily[actual_daily$date >=begin & actual_daily$date<= end,]
  exe_0 <- execution_daily[execution_daily$date >=begin & execution_daily$date<= end,] 
  minact<-min(actual_0[actual_0$sum>0,]$sum)
  minexc<-min(exe_0[exe_0$sum>0,]$sum)
  for (i in 1:320){
    if(!take_log){
      actual <- actual_0[actual_0$NEWID==i,5] 
      exe <- exe_0[exe_0$NEWID==i,5]
      acct_ave[i]<-mean(unlist(actual))
      exec_ave[i]<-mean(unlist(exe))
      dis_ac_exe[i] <- DTWDistance(actual,exe, window.type="sakoechiba",window.size=1)
     }
    else{
      logactual<-log(actual+minact)
      logexectut<-log(exe+minexc)  
      acct_ave[i]<-mean(unlist(logactual))
      exec_ave[i]<-mean(unlist(logexectut))
      dis_ac_exe[i] <- DTWDistance(logactual,logexectut, window.type="sakoechiba", window.size=1)
  }
  }
return(data.frame(actual=acct_ave, executed=exec_ave,distance=dis_ac_exe))
}


# output distance table 320*4 of (act & exe) or (log(act) & log (exe))
dis_ac_exe_final2 <-exe_actual_distance("2021-03-29","2021-07-04")
dis_ac_exe_final2 <- data.frame(NEWID=c(1:320),dis_ac_exe_final2)

dis_ac_exe_final_log <-exe_actual_distance("2021-03-29","2021-07-04",TRUE)
dis_ac_exe_final_log <- data.frame(NEWID=c(1:320),dis_ac_exe_final_log)

## make plot
par(mfrow=c(1,2))
plot(log(dis_ac_exe_final2$actual), log(dis_ac_exe_final2$distance))
plot(log(dis_ac_exe_final2$executed), log(dis_ac_exe_final2$distance))

ggplot(dis_ac_exe_final2)+geom_point()+
  aes(x=(executed), y=(actual), color=log(distance))+geom_abline(slope=1, intercept=0)+geom_quantile()+scale_x_log10()+scale_y_log10()

ggplot(dis_ac_exe_final2)+geom_point()+
  aes(x=(executed/actual), y=(distance))+geom_abline(slope=1, intercept=0)+geom_quantile()+scale_x_log10()+scale_y_log10()

```
log make less influence of gap between exe & act.
From

## group warehouses based on distance between effiency and actual credit
```{r}
##plot
scatter <- ggplot(dis_ac_exe_final)+
  geom_point(aes(x=NEWID,y=dis_ac_exe_final))+
  ylim(0,5000)
ggplotly(scatter)
```
From scatter plot, NEWID=9 has extermely high distance, over 4700, which means it's effiency is not stable from 2021-03-29 to 2021-07-04.

```{r}
## drop NEWID = 9 (new dataset:df)
df_ac_exe <- dis_ac_exe_final[-9,]
df_ac_exe.scaled <- scale(df_ac_exe[,-1])

dis_ac_exe_final2<-exe_actual_distance_2("2021-03-29","2021-07-04")
dis_ac_exe_final2 <- data.frame(NEWID=c(1:320),dis_ac_exe_final2)
df_ac_exe2 <- dis_ac_exe_final2[,-c(2,3)]
df_ac_exe.scaled2 <- scale(df_ac_exe2[,-1])

## choose best K
fviz_nbclust(df_ac_exe,kmeans, method = "wss") +
geom_vline(xintercept = 3, linetype = 2)   ## Best K = 3

fviz_nbclust(df_ac_exe2,kmeans, method = "wss") + geom_vline(xintercept = 3, linetype = 2) 
## Compute k-means with k = 3
set.seed(1)
kmdis_ac_exe <- kmeans(df_ac_exe.scaled, centers = 3, nstart = 25)
print(kmdis_ac_exe)

kmdis_ac_exe2 <- kmeans(df_ac_exe.scaled2, centers = 3, nstart = 25)

## Visualize k-means
fviz_cluster(kmdis_ac_exe, df_ac_exe, geom="text") ##use it

fviz_cluster(kmdis_ac_exe2, df_ac_exe2, geom="text") ##use it

## add clutering info to original dataset
df_ac_exe_cluster <- cbind(df_ac_exe, cluster = kmdis_ac_exe$cluster)
df_ac_exe_cluster[nrow(df_ac_exe_cluster)+1,] <- dis_ac_exe_final[9,]
df_ac_exe_cluster[,3][df_ac_exe_cluster[,3]==9] <- NA
df_ac_exe_cluster <- arrange(df_ac_exe_cluster,NEWID)
```

## Validation
```{r}
distanceAcExe <- dis_ac_exe_final[dis_ac_exe_final$NEWID %in% c(202,236,241,274,288),]
distanceAcExe$logDis <- log(distanceAcExe$dis_ac_exe_final)

log_effi <- efficiency_daily[efficiency_daily$date >="2021-03-29" & efficiency_daily$date<= "2021-07-04",]

join <- left_join(log_effi,dis_ac_exe_final,by="NEWID")
join_test <- join[join$NEWID %in% c(274,241,288,236,202),]

plot1 <- ggplot()+
  geom_boxplot(data=join_test,aes(x=factor(NEWID),y=efficiency),color="blue",notch=TRUE)

plot1+
  geom_point(data=distanceAcExe,aes(x=factor(NEWID),y=logDis),color="red")

test <- actual_daily[actual_daily$NEWID==274,]
test <- test[test$date >="2021-03-29" & test$date<= "2021-07-04",]

testEXE <- execution_daily[execution_daily$NEWID==274,]
testEXE <- testEXE[testEXE$date >="2021-03-29" & testEXE$date<= "2021-07-04",]

testEFFI <- efficiency_daily[efficiency_daily$NEWID==274,]
testEFFI <- testEFFI[testEFFI$date >="2021-03-29" & testEFFI$date<= "2021-07-04",]

log(dis_ac_exe_final[dis_ac_exe_final$NEWID==274,2])

ggplot(testEXE)+
  geom_line(aes(x=date,y=sum))+
  geom_hline(yintercept = 6.820545)

```
make a function to pick 5 of each group, show their actual series, exe, efficiency, show difference, interpret it.

```{r}
exe_actual_validate <- function(begin,end,ID1,ID2,ID3,ID4,ID5){
  # begin <- "2021-03-29"
  # end <- "2021-07-04"
  # ID1 <- 274
  # ID2 <- 9
  ac_validate <- actual_daily[actual_daily$NEWID %in% c(ID1,ID2,ID3,ID4,ID5),]
  ac_validate <- ac_validate[ac_validate$date >=begin & ac_validate$date<= end,]
  exe_validate <- execution_daily[execution_daily$NEWID %in% c(ID1,ID2,ID3,ID4,ID5),]
  exe_validate <- exe_validate[exe_validate$date >=begin & exe_validate$date<= end,]
  acplot <- ggplot(ac_validate) + 
    geom_line(aes(x=date,y=sum,color=factor(NEWID)))+
    ylab("actual_credit")
  exeplot <- ggplot(exe_validate) +
    geom_line(aes(x=date,y=sum,color=factor(NEWID)))+
    ylab("execution_credit")
  # newlist <- 
  return(list(acplot,exeplot))
}

exe_actual_validate("2020-11-29","2021-01-24",98,232,34,274,288)
```

```{r}
exe_actual_validate_one_warehouse <- function(begin,end,ID1){
  # begin <- "2021-03-29"
  # end <- "2021-07-04"
  # ID1 <- 274
  # ID2 <- 9
  ac_validate <- actual_daily[actual_daily$NEWID %in% c(ID1),]
  ac_validate <- ac_validate[ac_validate$date >=begin & ac_validate$date<= end,]
  exe_validate <- execution_daily[execution_daily$NEWID %in% c(ID1),]
  exe_validate <- exe_validate[exe_validate$date >=begin & exe_validate$date<= end,]
  datatemp<- data.frame(date= exe_validate$date, 
                        actual=ac_validate$sum, execute=exe_validate$sum)
  acplot <- ggplot(reshape2::melt(datatemp, id="date")) + 
    geom_line()+aes(x=date,y=value,color=variable)+
    ylab("credit")+ggtitle(paste("warehouse:",ID1))
  # newlist <- 
  return(acplot)
}

date.start="2021-03-29"
date.end="2021-07-04"
gridExtra::grid.arrange(exe_actual_validate_one_warehouse(date.start,date.end,98),
                        exe_actual_validate_one_warehouse(date.start,date.end,232),
                        exe_actual_validate_one_warehouse(date.start,date.end,34),
                        exe_actual_validate_one_warehouse(date.start,date.end,274),
                        exe_actual_validate_one_warehouse(date.start,date.end,288),
                        ncol=5)
```

```{r}
library(pheatmap)
pheatmap(log(dis_ac_exe_final2[,-1]+1))
set.seed(1)
fitres <- pheatmap(log(dis_ac_exe_final2[,-1]+1),kmeans_k = 6)
fitres$kmeans$cluster

# cluster 2--high distance & big actual
exe_actual_validate_one_warehouse(date.start,date.end,1)
# cluster1--fairly high dis & big actual
exe_actual_validate_one_warehouse(date.start,date.end,118)
# cluster3--relatively high & big
exe_actual_validate_one_warehouse(date.start,date.end,273)

# cluster5--low & very low
exe_actual_validate_one_warehouse(date.start,date.end,2)
```

